upwell_IDX

### 2020-04-21 upwell_IDX.Rmd
update

```{r}
# Installing and loading libraries

library(circular)
## devtools::install_github("robwschlegel/coastR")
library(gridExtra)
library(geosphere)
library(tidyverse)
library(heatwaveR)
library(coastR)
```

# Steps
  - Use the circular function
  - Determine daily wind data obtained from 3hr intervals

```{r, eval=FALSE}
wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
                     col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))

wind_2 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))

wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
                     col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))

# Selecting the important columns for each of the datasets
## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
wind_fix <- function(df){
wind <- df %>%
  mutate(date = as.Date(as.character(date)),
         hour = as.numeric(as.character(hour)),
         dir = as.numeric(as.character(dir)),
         speed = as.numeric(as.character(speed)),
         temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
                                station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
                                station_name  == "PORT NOLLOTH" ~"Port Nolloth",
                                station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
                                station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
                                station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay")) %>% 
  select(temp_sites, date, hour, dir, speed) %>% 
  drop_na()
}

wind_fix_1 <- wind_fix(df = wind_1)
wind_fix_2 <- wind_fix(df = wind_2)
wind_fix_3 <- wind_fix(df = wind_3)

# Combine and save
wind_data <- rbind(wind_fix_3, wind_fix_2, wind_fix_1)
save(wind_data, file = "Data/wind_data.RData")

load("Data/wind_data.RData")
selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Hout Bay")

# RWS: Please figure out why some wind direction values are returned as NA
# RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
# RWS: Please look through these resources and make corrections accordingly:
# https://cran.r-project.org/web/packages/rWind/rWind.pdf
# https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
wind_daily <- wind_data %>%
  # ungroup() %>%
  # mutate(date = as.Date(date))%>% # Changing the date to date format
  filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
  dplyr::group_by(temp_sites, date) %>%
  dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
                                                           template = 'geographics')),2), # Direction using the circular function
                   mean_speed = round(mean(speed),2)) # Calculating the mean
save(wind_daily, file = "Data/wind_daily.RData") 
```

@@

```{r}
load("Data/wind_daily.RData") 
wind_daily <- wind_daily %>% 
  dplyr::rename(sites = temp_sites)
load("Data/site_list_sub.Rdata")
```

@@

# Plotting function
plot_sites <- function(site_list, buffer, dist){

  # Find the point 200 km from the site manually to pass to ggplot
  heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
                                              b = site_list$heading, d = dist))

  # Add the new coordinates tot he site list
  site_list <- site_list %>% 
    mutate(lon_dest = heading2$lon,
           lat_dest = heading2$lat)

  # Visualise
  world_map +
    geom_segment(data = site_list, colour = "red4", 
                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
                             max(site_list$lon + buffer)),
                    ylim = c(min(site_list$lat - buffer), 
                             max(site_list$lat + buffer))) +
    labs(x = "", y = "", colour = "Site\norder")
}
south_africa_away <- plot_sites(south_africa_away_wide, 1, 100000)
south_africa_away
```

@@

```{r}
UI_angle <- wind_daily %>%  # Making reference to the wind daily data created above
  dplyr::rename(site = sites) %>%  # Renaming sites to site
  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to South Africa wide data column 2 and 21. by site 
  drop_na() %>% # removing the na values
  dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
  dplyr::rename(t = date) #renaming date to t

@@

```{r}
# Loading the insitu temperature data along the wc
load("Data/SACTN_US.RData") # temperature data for all the sites with a 30year time series
# exceedance <-
#   function(data,
#            x = t,
#            y = temp,
#            threshold,
#            below = FALSE,
#            minDuration = 5,
#            joinAcrossGaps = TRUE,
#            maxGap = 2,
#            maxPadLength = 3)
# Upwelling temperature threshold 12.4 / 30th percentile
SACTN_upwell <- UI_angle %>% 
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t), 
  #        max_t = max(t)) %>% 
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")), # creating a column of climaatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>% 
  select(-data, -clim) %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance) # selecting only these variables
# Calculate quantiles of upwelling index

>>>>

```{r}
# # Installing and loading libraries
# 
# library(circular)
# ## devtools::install_github("robwschlegel/coastR")
# library(gridExtra)
# library(geosphere)
# library(tidyverse)
# library(heatwaveR)
# library(coastR)
# ```
# 
# # Steps
#   - Use the circular function
#   - Determine daily wind data obtained from 3hr intervals
# 
# ```{r, eval=FALSE}
# wind_1 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub", "speed", "dir"))
# 
# wind_2 <- read.delim("Documents/Data/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
# 
# wind_3 <- read.delim("Data/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings = "",
#                      col.names = c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir"))
# 
# # Selecting the important columns for each of the datasets
# ## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
# ## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected
# wind_fix <- function(df){
# wind <- df %>%
#   mutate(date = as.Date(as.character(date)),
#          hour = as.numeric(as.character(hour)),
#          dir = as.numeric(as.character(dir)),
#          speed = as.numeric(as.character(speed)),
#          temp_sites = case_when(station_name == "CAPE TOWN TABLE BAY" ~ "Sea Point",
#                                 station_name == "CAPE TOWN - ROYAL YACHT CLUB" ~ "Sea Point",
#                                 station_name  == "PORT NOLLOTH" ~"Port Nolloth",
#                                 station_name  == "CAPE TOWN SLANGKOP" ~ "Hout Bay",
#                                 station_name  == "LAMBERTSBAAI NORTIER" ~ "Lamberts Bay",
#                                 station_name  == "LANGEBAANWEG AWS" ~ "Saldanha Bay")) %>% 
#   dplyr::select(temp_sites, date, hour, dir, speed) %>% 
#   drop_na()
# }
# 
# wind_fix_1 <- wind_fix(df = wind_1)
# wind_fix_2 <- wind_fix(df = wind_2)
# wind_fix_3 <- wind_fix(df = wind_3)
# 
# # Combine and save
# wind_data <- rbind(wind_fix_3, wind_fix_2, wind_fix_1)
# # save(wind_data, file = "Data/wind_data.RData")
# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay") #Removed Hout Bay as this is not one of my sites.
# 
# # RWS: Please figure out why some wind direction values are returned as NA
# # RWS: Also, it's not enough to just calculate mean speed. It needs to be taken into account with the directions.
# # RWS: Please look through these resources and make corrections accordingly:
# # https://cran.r-project.org/web/packages/rWind/rWind.pdf
# # https://www.kaggle.com/srlightfoote/example-wind-analysis-in-r
# 
# # Errors in the wind data for Port Nolloth - Recieved the correct wind data with no missing hours
# 
# wind_daily <- wind_data %>%
#   # ungroup() %>%
#   # mutate(date = as.Date(date))%>% # Changing the date to date format
#   filter(temp_sites %in% selected_sites) %>%  # Matching the sites to wind temp sites
#   dplyr::group_by(temp_sites, date) %>%
#   dplyr::summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees", 
#                                                            template = 'geographics')),2)) #, # Direction using the circular function
#                   # mean_speed = round(mean(speed),2)) # Calculating the mean
# save(wind_daily, file = "Data/wind_daily.RData") 
```

@@

```{r}
load("Data/wind_daily.RData") 
wind_daily <- wind_daily %>%
  dplyr::rename(sites = temp_sites) %>% 
  drop_na()
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
```

@@

# Plotting function
# plot_sites <- function(site_list, buffer, dist){

  # Find the point 200 km from the site manually to pass to ggplot
#   heading2 <- data.frame(geosphere::destPoint(p = select(site_list, lon, lat),  
#                                               b = site_list$heading, d = dist))
#   
#   # Add the new coordinates tot he site list
#   site_list <- site_list %>% 
#     mutate(lon_dest = heading2$lon,
#            lat_dest = heading2$lat)
#   
#   # Visualise
#   world_map +
#     geom_segment(data = site_list, colour = "red4", 
#                  aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
#     geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
#     geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
#     coord_cartesian(xlim = c(min(site_list$lon - buffer), 
#                              max(site_list$lon + buffer)),
#                     ylim = c(min(site_list$lat - buffer), 
#                              max(site_list$lat + buffer))) +
#     labs(x = "", y = "", colour = "Site\norder")
# }
# south_africa_away <- plot_sites(south_africa_away_wide, 1, 100000)
# south_africa_away
```

@@

```{r}
UI_angle <- wind_daily %>%  # Making reference to the wind daily data created above
  dplyr::rename(site = sites) %>%  # Renaming sites to site
  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>%  # Joining wind data to South Africa wide data column 2 and 21. by site 
  drop_na() %>% # removing the na values
  dplyr::rename(coast_angle = heading) %>% # Renaming the column heading to coast_angle
  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # applying the ui formula
  dplyr::rename(t = date) %>% 
  drop_na()#renaming date to t

@@

```{r}
# Loading the insitu temperature data along the wc
load("Data/SACTN_US.RData") # temperature data for all the sites with a 30year time series
# exceedance <-
#   function(data,
#            x = t,
#            y = temp,
#            threshold,
#            below = FALSE,
#            minDuration = 5,
#            joinAcrossGaps = TRUE,
#            maxGap = 2,
#            maxPadLength = 3)
# Upwelling temperature threshold 12.4 / 30th percentile
SACTN_upwell <- UI_angle %>% 
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t), 
  #        max_t = max(t)) %>% 
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1992-01-01", "2016-12-31")), # creating a column of climaatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>% 
  select(-data, -clim) %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance) # selecting only these variables
# Calculate quantiles of upwelling index
